Bayesian beta regressions with brms in R: A tutorial for phoneticians

### 2025-02-10 README.qmd
render all

The HTML version of the paper can be found [here](manuscript/manuscript.html).

>>>>

The code of the manuscript is in `manuscript/manuscript.qmd`. The `code` folder contains code with data checks and plotting.

The rendered HTML version of the paper and the data checks can be found on the [repository website](https://stefanocoretta.github.io/beta-phon/):

- [manuscript](manuscript/manuscript.html).
- [data checks](code/01_data_checks.html).

### 2025-02-10 manuscript/manuscript.qmd
render all

---
title: A tutorial of Bayesian beta regressions with brms in R
author:
  - name: Stefano Coretta
    email: s.coretta@ed.ac.uk
    affiliations: 
        - id: uoe
          name: University of Edinburgh
          department: Linguistics and English Language
          address: 3 George Sq
          city: Edinburgh
          state: United Kingdom
          postal-code: EH8 9AD
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Paul Bürkner
    email: paul.buerkner@gmail.com
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: 
  - linguistics.bib
  - statistics.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
execute:
  cache: true
---

@@

While most researchers approach proportions with Gaussian regression models, proportion are not Gaussian by nature, since they are continuous variables bounded between 0 and 1. Thus, regression models with proportions as outcome variable should be fitted using a likelihood family that assumes such unit interval data. A common and flexible distribution for this purpose is the beta distribution. This tutorial introduces researchers to beta regression models in R using the package brms. Familiarity with regression modelling in R with a package like lme4 is assumed, but no prior knowledge of Bayesian statistics is necessary. The rest of the paper is structured as follows: @sec-beta introduces the mathematical underpinnings of the beta distribution, @sec-bayes provides the readers with a brief explanation and justification of Bayesian regression models, while @sec-case-1 and @sec-case-2 illustrate how to fit Bayesian beta regression models with two case studies.

>>>>

---
title: "Bayesian beta regressions with brms in R: A tutorial for phoneticians"
author:
  - name: Stefano Coretta
    email: s.coretta@ed.ac.uk
    affiliations: 
        - id: uoe
          name: University of Edinburgh
          department: Linguistics and English Language
          address: 3 George Sq
          city: Edinburgh, United Kingdom
          postal-code: EH8 9AD
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Paul Bürkner
    email: paul.buerkner@gmail.com
    affiliations:
        - id: dort
          name: TU Dortmund University
          department: Department of Statistics
          adress: Technische Universität Dortmund
          city: Dortmund, Germany
          postal-code: 44221
abstract: |
  Phonetic research frequently involves analyzing numeric continuous outcome variables, such as durations, frequencies, loudness, and ratios. Another commonly used outcome type is proportions, including measures like the proportion of voicing during closure, gesture amplitude, and nasalance. Despite their bounded nature, proportions are often modeled using Gaussian regression, largely due to the default settings of commonly used statistical functions in R (e.g., lm() and lmer() from lme4). This practice persists in teaching and research, despite the fact that Gaussian models assume unbounded continuous data and may poorly fit proportion data. To address this issue, this tutorial introduces beta regression models, a more appropriate statistical approach for analyzing proportions. The beta distribution provides a flexible framework for modelling continuous data constrained between 0 and 1. The tutorial employs the brms package in R and assumes familiarity with regression modeling but no prior knowledge of Bayesian statistics. The tutorial includes two case studies illustrating the practical implementation of Bayesian beta regression models. Data and code are available at <https://github.com/stefanocoretta/beta-phon>.
keywords: 
  - acoustics
  - articulation
  - proportions
  - Bayesian regressions
  - beta distribution
date: last-modified
bibliography: 
  - linguistics.bib
  - statistics.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
  elsevier-html: default
execute:
  cache: true
---

@@

While most researchers approach proportions with Gaussian regression models, proportion are not Gaussian by nature, since they are continuous variables bounded between 0 and 1. Thus, regression models with proportions as outcome variable should be fitted using a likelihood family that assumes such unit interval data. A common and flexible distribution for this purpose is the beta distribution. This tutorial introduces researchers to beta regression models in R using the package brms. Familiarity with regression modelling in R with a package like lme4 is assumed, but no prior knowledge of Bayesian statistics is necessary. The rest of the paper is structured as follows: @sec-beta introduces the mathematical underpinnings of the beta distribution, @sec-bayes provides the readers with a brief explanation and justification of Bayesian regression models, while @sec-case-1 and @sec-case-2 illustrate how to fit Bayesian beta regression models with two case studies. The research compendium with data and code can be found at <https://github.com/stefanocoretta/beta-phon>.

### 2025-02-10 manuscript/manuscript.tex
render all

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Bayesian beta regressions with brms in R: A tutorial for phoneticians},
  pdfauthor={Stefano Coretta; Paul Bürkner},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Bayesian beta regressions with brms in R: A tutorial for
phoneticians}
\author[1]{Stefano Coretta%
\corref{cor1}%
\fnref{fn1}}
 \ead{s.coretta@ed.ac.uk} 
\author[]{Paul Bürkner%
%
}
 \ead{paul.buerkner@gmail.com}

\affiliation[1]{organization={University of Edinburgh, Linguistics and
English Language},addressline={3 George
Sq},city={Edinburgh},postcode={EH8 9AD},postcodesep={}}

@@

\begin{abstract}
This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing
elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget
velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris
sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta
egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet
auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo
molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae.
Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at
tempus elit.
\end{abstract}

\begin{keyword}
    keyword1 \sep 
    keyword2
\end{keyword}
\end{frontmatter}

@@

While most researchers approach proportions with Gaussian regression
models, proportion are not Gaussian by nature, since they are continuous
variables bounded between 0 and 1. Thus, regression models with
proportions as outcome variable should be fitted using a likelihood
family that assumes such unit interval data. A common and flexible
distribution for this purpose is the beta distribution. This tutorial
introduces researchers to beta regression models in R using the package
brms. Familiarity with regression modelling in R with a package like
lme4 is assumed, but no prior knowledge of Bayesian statistics is
necessary. The rest of the paper is structured as follows:
Section~\ref{sec-beta} introduces the mathematical underpinnings of the
beta distribution, Section~\ref{sec-bayes} provides the readers with a
brief explanation and justification of Bayesian regression models, while
Section~\ref{sec-case-1} and Section~\ref{sec-case-2} illustrate how to
fit Bayesian beta regression models with two case studies.

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Bayesian beta regressions with brms in R: A tutorial for phoneticians},
  pdfauthor={Stefano Coretta; Paul Bürkner},
  pdfkeywords={acoustics, articulation, proportions, Bayesian
regressions, beta distribution},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Bayesian beta regressions with brms in R: A tutorial for
phoneticians}
\author[1]{Stefano Coretta%
\corref{cor1}%
\fnref{fn1}}
 \ead{s.coretta@ed.ac.uk} 
\author[2]{Paul Bürkner%
%
}
 \ead{paul.buerkner@gmail.com}

\affiliation[1]{organization={University of Edinburgh, Linguistics and
English Language},addressline={3 George Sq},city={Edinburgh, United
Kingdom},postcode={EH8 9AD},postcodesep={}}
\affiliation[2]{organization={TU Dortmund University, Department of
Statistics},city={Dortmund, Germany},postcode={44221},postcodesep={}}

@@

\begin{abstract}
Phonetic research frequently involves analyzing numeric continuous
outcome variables, such as durations, frequencies, loudness, and ratios.
Another commonly used outcome type is proportions, including measures
like the proportion of voicing during closure, gesture amplitude, and
nasalance. Despite their bounded nature, proportions are often modeled
using Gaussian regression, largely due to the default settings of
commonly used statistical functions in R (e.g., lm() and lmer() from
lme4). This practice persists in teaching and research, despite the fact
that Gaussian models assume unbounded continuous data and may poorly fit
proportion data. To address this issue, this tutorial introduces beta
regression models, a more appropriate statistical approach for analyzing
proportions. The beta distribution provides a flexible framework for
modelling continuous data constrained between 0 and 1. The tutorial
employs the brms package in R and assumes familiarity with regression
modeling but no prior knowledge of Bayesian statistics. The tutorial
includes two case studies illustrating the practical implementation of
Bayesian beta regression models. Data and code are available at
\url{https://github.com/stefanocoretta/beta-phon}.
\end{abstract}

\begin{keyword}
    acoustics \sep articulation \sep proportions \sep Bayesian
regressions \sep 
    beta distribution
\end{keyword}
\end{frontmatter}

@@

While most researchers approach proportions with Gaussian regression
models, proportion are not Gaussian by nature, since they are continuous
variables bounded between 0 and 1. Thus, regression models with
proportions as outcome variable should be fitted using a likelihood
family that assumes such unit interval data. A common and flexible
distribution for this purpose is the beta distribution. This tutorial
introduces researchers to beta regression models in R using the package
brms. Familiarity with regression modelling in R with a package like
lme4 is assumed, but no prior knowledge of Bayesian statistics is
necessary. The rest of the paper is structured as follows:
Section~\ref{sec-beta} introduces the mathematical underpinnings of the
beta distribution, Section~\ref{sec-bayes} provides the readers with a
brief explanation and justification of Bayesian regression models, while
Section~\ref{sec-case-1} and Section~\ref{sec-case-2} illustrate how to
fit Bayesian beta regression models with two case studies. The research
compendium with data and code can be found at
\url{https://github.com/stefanocoretta/beta-phon}.
